---
title: "dplyr Verbs"
output: html_notebook
---


```{r message=FALSE}
library(dplyr)
library(readr)
```





# Read  dataset
```{r}
base_data<-read_csv("../../rawdata/consumo_db.csv",col_types = cols())
base_data
```

There are several commonly used verbs in the dplyr package that are useful for manipulating and summarizing data. Here are some of the most common verbs in dplyr:

# filter(): This verb is used to select rows of a data frame based on certain conditions.


```{r}
base_data %>% filter(SalesOrder=="34046361")
```



# select(): This verb is used to select specific columns of a data frame.


```{r}
base_data %>% select(SalesOrder,UM)
```


# group_by(): This verb is used to group the rows of a data frame by one or more columns.


```{r}
base_data %>% group_by(EnduserCode) %>% count()
```


# arrange(): This verb is used to sort the rows of a data frame by one or more columns.

```{r}
base_data %>% group_by(EnduserCode) %>% count() %>% arrange(n %>% desc())
```


# mutate(): This verb is used to create new columns in a data frame based on existing columns.

```{r}
base_data<-base_data %>% mutate(InvoiceQty=as.numeric(InvoiceQty))
```


# mutate() Convert to date type

```{r}
base_data<-base_data %>% mutate( date = lubridate::dmy(InvoiceDate))
```
## Plot

```{r fig.height=4, fig.width=10}
base_data %>% filter(!EnduserCode %>% is.na())

base_data %>% filter(EnduserCode == "Client 002") %>% select(date,ShipQty) %>% tidyr::drop_na() %>% group_by(date) %>% 
  summarise(ShipQty=sum(ShipQty)) %>%
  ggplot()+
  geom_line(aes(x=date,y=ShipQty),color='red')+
  geom_point(aes(x=date,y=ShipQty),color='orange')+
  
  theme_bw()
```




# summarize(): This verb is used to calculate summary statistics of a data frame, such as mean, median, or total.

```{r}
base_data %>% group_by(EnduserCode) %>% summarise(qty=sum(InvoiceQty,na.rm = T))
```


# OK, let's build a model...


**Q1: Can we predict the shipQty considering : **
                   * "Region",
                   * "Process",
                   * "StringGroup",
                   * "WellName"

## Aggregate ShipQty by User and SalesOrder
```{r}
aggregated_salesorder_data<-base_data %>% group_by(EnduserCode,SalesOrder) %>% summarise(sales_order_num=n(),
                                                                                         sum_Shipqty=sum(ShipQty)) 
aggregated_salesorder_data
```

## Select only SalesOrders with same data in the shipment


```{r}
select_sales_orders<-base_data %>% 
  select(SalesOrder,Region,Process,StringGroup,WellName) %>% 
  unique() %>% mutate(SalesOrder=as.factor(SalesOrder)) %>% 
  group_by(SalesOrder) %>% 
  summarise(n=n()) %>% filter(n==1) %>% 
  select(SalesOrder)%>% unlist() %>% 
  unname()

aggregated_salesorder_data <- aggregated_salesorder_data %>% 
  filter( SalesOrder %in% select_sales_orders)
```



```{r}
aggregated_salesorder_data <- left_join(aggregated_salesorder_data, base_data %>%
                                        select(SalesOrder,Region,Process,StringGroup,WellName) %>% 
                                          unique(), by="SalesOrder")
```

## Split well info to reduce the levels of the column
```{r}
aggregated_salesorder_data<-aggregated_salesorder_data %>% 
  mutate(WellName=stringr::str_split(WellName,pattern = " ")[[1]][1]) 
```

# Split train and test
```{r}
library(caret)
trainset <- aggregated_salesorder_data %>% ungroup() %>% na.omit() %>% sample_frac(0.7)
testset <- setdiff(aggregated_salesorder_data %>% ungroup(),trainset)
```

# Train simple Model

```{r message=FALSE}
library(caret)
library(doMC)
registerDoMC(cores = 4)
selected_variables<-c(#"sales_order_num",
                   "Region",
                   "Process",
                   "StringGroup",
                   "WellName"
                   )
ctrl_fast <- trainControl(
  method = "repeatedcv",
  repeats = 2,
  number = 5,
  returnResamp = 'final',
  savePredictions = 'final',
  verboseIter = F,
  allowParallel = T
)


cartFit <- caret::train(
  x = trainset %>%  select(selected_variables),
  y = trainset %>%  select(sum_Shipqty) %>% unname() %>% unlist(),
  method = "rf",
  tuneLength=3,
  #verbose = 2,
  trControl = ctrl_fast,
  ntree = 500
)
```

# Model results
```{r}
cartFit$results
cartFit$finalModel
```

# Resample (CV) results
```{r}
cartFit$results %>%
  ggplot(aes(x=mtry,y=Rsquared))+
  geom_point(color = 'red') +
  geom_errorbar(aes(ymin=Rsquared-RsquaredSD,ymax=Rsquared+RsquaredSD),width=0.2)+
  theme_classic()
```

# Test Results
```{r}
testset <- testset %>% na.omit()
predictions<-predict(cartFit,testset)
predictions<-data.frame(pred=predictions,obs=testset$sum_Shipqty %>% na.omit())
Rsquared<-postResample(pred=predictions$pred,obs=predictions$obs)[[2]]
predictions %>% 
  ggplot(aes(pred,obs))+
  geom_point(shape=21,alpha=0.5)+
  #geom_label(y=300,x=10,label=paste("R^2:",round(Rsquared,digits = 4)),color="blue",size=4,)+
  labs(title="Random Forest for Aggregated shpQty", subtitle="Predicted vs. Observed")+
  geom_abline(intercept = 0,
              slope = 1,
              color = "red",
              size = 1)+
  theme_bw()
```